An Empirical Analysis of the Role of Energy in Economic Growth

### 2013-08-02 EconGrowth2.Rnw
Added section on LINEX parameter estimation.

Each of the economic growth models discussed in 
Section~\ref{sec:Prod_Func_Forms} has parameters which must be estimated (fitted)
for each economy (see Section~\ref{sec:Economy_Types}) and 
each energy quantification (see Section~\ref{sec:Quantification_of_Energy}) 
via least-squares methods. We used \texttt{R} \citep{R}, specifically the \texttt{nls} and 
\texttt{cesEst} \citep{Henningsen:2011td} functions for the parameter estimation process.

The following subsections discuss 
parameter estimation methods and results.

@@

\noindent If none of the three reparameterizations converged, 
the parameterization with the smallest sum of square errors ($SSE$)
was used. \ref{sec:Convergence_Resample_Models} shows the 
fraction of unconverged models for each model type. 
The above approach resulted in very few unconverged
Cobb-Douglas resample models, even with \Sexpr{numResamples()} resamples.

@@

\begin{quote}
\ldots the Levenberg-Marquardt and the PORT algorithms 
are---at least in this study---most likely to find the 
coefficients that give the best fit to the model, 
where the PORT algorithm can be used to restrict the estimates 
to the economically meaningful region.
\end{quote}

In our testing, we found that to be mostly true. 
\texttt{PORT} nearly always provided lower
$sse$ than \texttt{L-BFGS-B}, despite the fact that 
\texttt{L-BFGS-B} often reports convergence
and \texttt{PORT} does not for the same economic data.

@@

@@

Grid search etc. (Matt, Randy to populate)

This section should cover both technique and results. 
Organize along coordinates of analysis.

1. Discuss parameter estimation for each model here.

CES Model: It is possible for L-BFGS-B to report convergence and PORT to report non-convergence 
but nevertheless reach a better solution (smaller sse).

2. Show scatter plot results.

>>>>

Each of the economic growth models discussed in 
Section~\ref{sec:Prod_Func_Forms} has parameters which must be estimated (fitted)
for each economy (see Section~\ref{sec:Economy_Types}) and 
each energy quantification (see Section~\ref{sec:Quantification_of_Energy}) 
via least-squares methods. We used \texttt{R} \citep{R} functions
\texttt{nls} and \texttt{cesEst} \citep{Henningsen:2011td} 
for that purpose. The following subsections discuss 
parameter estimation methods and results.

@@

\noindent If none of the three reparameterizations converged, 
the parameterization with the smallest sum of square errors ($sse$)
was used. \ref{sec:Convergence_Resample_Models} shows the 
fraction of unconverged models for each model type. 
The above approach resulted in very few unconverged
Cobb-Douglas resample models, even with \Sexpr{numResamples()} resamples.

@@

\begin{quote}
\ldots the Levenberg-Marquardt and the \texttt{PORT} algorithms 
are---at least in this study---most likely to find the 
coefficients that give the best fit to the model, 
where the PORT algorithm can be used to restrict the estimates 
to the economically meaningful region.
\end{quote}

In our testing, we found that to be mostly true. 
\texttt{PORT} nearly always provided lower
$sse$ than \texttt{L-BFGS-B}, despite the fact that 
\texttt{L-BFGS-B} often reports convergence
and \texttt{PORT} does not for the same data
(i.e., for the same $y$, $k$, $l$, and $e$ data).

@@

%++++++++++++++++++++++++++++++
\subsection{LINEX Models}
\label{sec:LINEX_Parameter_Estimation}
%++++++++++++++++++++++++++++++

The LINEX model is given in Equation~\ref{eq:LINEX}. 
If the fitting algorithm drives $a_0 \rightarrow 0$, 
any changes to $c_t$ have no effect on $sse$, 
leading the fitting algorithm to report non-convergence. 
To avoid this problem, we transformed the LINEX model into

\begin{equation} \label{eq:LINEX_transformed}
  y = Ae; 
  A = \mathrm{e}^{\left[2 a_0 \left(1 - \frac{1}{\rho_k} \right) 
    + a_1 \left(\vphantom{\frac{1}{\rho_k}}\rho_l - 1 \right)\right]},
\end{equation}

\noindent where $a_1 \equiv a_0 c_t$. 
We used the \texttt{nls} function in \texttt{R}
\citep{R} to estimate least-squares values for $a_0$ and $a_1$ in 
Equation~\ref{eq:LINEX_transformed}. We recovered
a value for $c_t$ with $c_t = a_1 / a_0$. The reparameterization
of Equation~\ref{eq:LINEX_transformed} reduced the number of 
unconverged resample fits to nearly zero.

**** Include LINEX scatter graphs here. Cover all 
quantifications of energy? CIs? TBD which to include. ****

Figure~\ref{fig:LINEX_GDP_Lattice_Graph} compares LINEX model predictions to historical data.

<<LINEX_GDP_Lattice_Graph, fig.pos="H", fig.width=ninePanelGraphWidth, fig.height=ninePanelGraphHeight, fig.cap="LINEX results.">>=
createLINEXLatticeGraph()
@

@@
